---
subtitle: "Project Proposal"
author: "Apache Junction Armchairs: Ellie, Ryan, Sude, Darren"
date: "2/27/2020"
output: pdf_document
---

```{r load-packages}
# include packages as needed
library(tidyverse)

```

```{r load-data}
# load the data
gentdata <- read_csv("data/gentdata.csv")
```

**ERASE ALL INSTRUCTIONS BEFORE SUBMITTING THE PROPOSAL**

### Project Description


- What is the research question you wish to explore? 

The research question that we wish to explore is what factors have contributed to the gentrification of the Research triangle (Counties including: Durham, Wake, Orange and Chatham) and where is it occuring most. 

*This includes an introduction to the subject matter you’re investigating, the motivation for the for your research question (citing any relevant literature), and your hypotheses regarding the research question of interest.*

In the past 10 years, gentrification has become a major issue in Durham. We are interested in finding out where and why gentrification is happening most. We predict that these variables and our analysis will give us insight into the displacement of the Black population in Durham and the entire Research Trianlge. We chose to look compare census tract data in Durham, Wake, Orange, and Chatham counties in 2010 and 2017 to analyze this movement. 

We hypothesize that census tracts in which there is an increase in college educated whites, an increase in median household price and an increase in the amount of people who obtain a bachelor degree or higher will correspond to a decrease in Black population. In addition to exploring the mechanisms behind this displacement, we are interested in investigating where this displacement occurs. We want to be able to pinpoint the neighborhoods with the highest rate of incoming college-educated white people and the highest rate of Black exodus.

- Describe the dataset you wish to explore. 
*This includes describing the observations in the data, a general description of the variables, and how the data was originally collected (not how you found the data but how the original curator of the data collected it.)*

We put together a dataset of relevant variables that we wish to explore from https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml. 

All of our demographic data is framed in terms of percentage point change from 2010-2017. By measuring our variables in percentage point changes, we are making our variable more robust to population changes in the years between 2010-2017. We are more focused on the proportion of demographic changes, as opposed to changes of a particular demographic group in a vacuum.

## The Data

- Use the glimpse function print a summary of the data frame at the end of your proposal.

```{r}
glimpse(gentdata)
```


