---
author: 'Apache Junction Armchairs: Ellie, Ryan, Sude, Darren'
date: "2/27/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: Project Proposal
---

```{r load-packages}
library(tidyverse)
```

```{r}
gentdata <- read_csv("data/gentdata.csv", col_names = TRUE, col_types = cols())
metadata <-read_csv("data/gentmetadata.csv", col_names = TRUE, col_types = cols())
```

  What factors have contributed to the gentrification of the Research triangle (counties including: Durham, Wake, Orange and Chatham), and where is it occuring most?
  
  In the past 10 years, development in Durham has skyrocketed, and the face of downtown has changed drastically. Alongside the rapid growth, gentrification has become a major issue in Durham. This 2018 New York Times sums up the changes in Durham well: "What has been largely overlooked is the cultural displacement that can accompany rapid urban change: the sense that home is not home anymore, at least for a portion of the population" (https://www.nytimes.com/2018/05/01/business/durham-real-estate-growth.html). We are interested in finding out where gentrification is happening, and why. We predict that the variables we have selected will give us insight into the displacement of the Black population in Durham and the rest of the Research Trianlge. We chose to compare census tract-level data in Durham, Wake, Orange, and Chatham counties in 2010 and 2017 to analyze this movement. 
  
   We hypothesize that census tracts in which there is an increase in college educated whites, an increase in median household price and an increase in the amount of people who obtain a bachelor degree or higher will correspond to a decrease in Black population. In addition to exploring the mechanisms behind this displacement, we are interested in investigating where this displacement occurs. We want to be able to pinpoint the neighborhoods with the highest rate of incoming college-educated white people and the highest rate of Black exodus.

  We put together a dataset of relevant variables from The Census Bureau's annual American Community Survey, which conducts a wide range of demographic surveys that covers almost everything a decentennial census covers. The data can be found here: https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml. 

  All of our demographic data is framed in terms of percentage point change from 2010-2017. By measuring our variables in percentage point changes, we are making our variable more robust to population changes in the years between 2010-2017. We are more focused on the proportion of demographic changes, as opposed to changes of a particular demographic group in a vacuum. Our "metadata" file explains what each variable is actually measuring. Median home price and median household income are the only two variables measured in monetary terms, as opposed to PP change.


```{r}
glimpse(gentdata)
glimpse(metadata)
```


